```{r}
# Load pacman and install required packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, caret, corrplot, psych, stringr, ggpubr, forecast, gridExtra, car, ggcorrplot)
options(scipen=999) # surpress scientific notation
```

```{r}
# Read dataset
bfat = read.csv("bodyfatmen.csv")
```

```{r}
# Fit model
model = lm(density ~ . , data = bfat)
```

```{r}
# Summary statistics
summary(model)
```
- Abdomen, followed by wrist, followed by neck and forearm are significant predictors, the rest not so much
- Adjusted R-squared value indicates a decent fit
- Good F-statistic indicates model is highly significant


## ===== Residuals analysis ===== ## 
```{r}
# Get studentized residuals
stud_res = rstudent(model)
```

```{r}
# Plot Residuals vs Fitted
ggplot(data.frame(Fitted = fitted(model), Residuals = stud_res), aes(x = Fitted, y = Residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals")
```
points randomly scattered around 0 => linearity
points evenly spread out across the fitted values => homoscedasticity (equal variance)
some points seem to be far apart => some outliers/influential points

## ===== End of Residuals analysis ===== ## 





## ===== Determining normality ===== ## 
```{r}
qqnorm(stud_res)
qqline(stud_res, col='red')
```
```{r}
qqPlot(model, main = "Q-Q Plot of Residuals")
```
qqplot indicates a rather normally distributed studentized residuals

```{r}
# Plot histogram
hist(stud_res, main = "Histogram of Residuals", xlab = "Residuals", ylab = "Frequency", breaks = 20)

# Impose Normal pdf
xpt = seq(min(stud_res), max(stud_res), by=0.1) # start, end, interval
n_den = dnorm(xpt, mean(stud_res), sd(stud_res)) # normal density w mean and sd of the data

# Get bin width from the histogram
hist_info <- hist(stud_res, breaks = 20, plot = FALSE)  # Store histogram data without plotting
bin_width <- hist_info$breaks[2] - hist_info$breaks[1]  # Compute actual bin width

ypt = n_den * length(stud_res) * bin_width # adjust the normal density to the length of the dataset * width of each bin
lines(xpt, ypt, col='blue')
```
histogram also indicates that the studentized residuals are normally distributed (with an outlier)

```{r}
shapiro.test(stud_res)
```

## ===== End of Determining normality ===== ## 





## ===== Identifying Leverage & Influence Points ===== ## 

```{r}
outliers <- which(abs(stud_res) > 3) 
# Reference for cutoff: https://online.stat.psu.edu/stat462/node/247/#:~:text=If%20an%20observation%20has%20a,than%203%20in%20absolute%20value.

print(outliers)
```
Only 1 outlier identified, at entry 96

```{r}
# Identifying Leverage points
# Reference: https://www.statology.org/leverage-in-r/
# Get hat values and store them in a data frame
hats <- data.frame(HatValue = hatvalues(model))

# Display sorted hat values (descending order)
hats[order(-hats$HatValue), , drop=FALSE]

# Implementing cutoff for leverage points using 2p/n
p = length(coef(model)) # number of predictors (including intercept)
n = nrow(bfat) # number of rows of data
leverage_cutoff = 2*p/n # 0.1111..
#print(leverage_cutoff)

# Show entries that exceed cutoff
which(hats$HatValue > leverage_cutoff)

```
- The leverage points are (5  31  36  39  41  42  54  86 106 159 175 206 216) based on the cutoff rule 2p/n
- On manual inspection of the points, nothing seem out of the ordinary

```{r}
# Identifying influence points using Cook's Distance 
# Cooks distance cutoff (F statistic - alpha, p, n-p)
alpha = 0.5
cook_cutoff <- qf(alpha, p, n - p) # calculate cutoff (p and n calculated above for leverage pts)
print(cook_cutoff)

# Plot cooks distance 
cooksd <- cooks.distance(model)
plot(cooksd, type = "h", main = "Cook’s Distance")
abline(h = cook_cutoff, col = "red")  # Threshold line

# Print influential points
high_influence <- which(cooksd > cook_cutoff)
print(high_influence)
```
Seems like there are no influential points based on Cook's Distance

```{r}
# Might want to look at DFFITS & DFBETAS for influence also

```

Even though there seem to be some outliers and leverage points, when all things considered and using Cook's Distance, none of them seem to be influential points which indicate that there may not be a need to remove any points. The outlier (96th record) is not a leverage point, and none of the leverage points are outliers.

## ===== End of Identifying Leverage & Influence Points ===== ## 





## ===== Multicollinearity ===== ## 

Next steps:
- Multicollinearity diagnosis
```{r}
# Correlation plot to check for pairwise correlations
cor_matrix <- cor(bfat[, -1])
# Plot correlation matrix
corrplot(cor_matrix, method = "color", type = "lower",
         tl.cex = 0.7,   # Reduce axis label size
         number.cex = 0.7,  # Reduce number size inside squares
         addCoef.col = "black",  # Show numbers inside squares
         title = "Correlation Plot of Predictors")
```

```{r}
# VIF analysis
vif_vals = vif(model)
print(vif_vals)
# Visualizing VIF
barplot(vif_vals, col = "skyblue", main = "Variance Inflation Factor (VIF)", las=2)
abline(h = 10, col = "red")
abline(h = 5, col = "blue")
```
- Using the cutoff rule of vif > 10, we see that 'weight', 'abdomen', and 'hip' are significant
- 'Chest' is borderline



